---
title: "Final Project R Reference"
author: "Julia Vineyard"
date: "12/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R Reference Guide {.tabset .tabset-pills}

## Loading Data and Packages

### here()
* here() lets you direct R to where you data is located in your directory

* To load here() you must use require(here) first

* Using here() first also tells you where R is currently pulling your data from

```{r warning= FALSE}
require(here)
```


### read.csv()

* Then you can combine here() with read.csv() to make a dataframe and bring it into the environment

* read.csv() allows R to read the data file that you are directing it to into a dataframe

* You can name the dataframe whatever you want

```{r warning= FALSE}
ginkgo= read.csv(here("data", "ginkgo_data_2021.csv"))
```

### library() 

* The library() function is another way to bring in a package, but only from the R directory

* You can name the dataframe whatever you want

```{r warning=FALSE}
palmer_penguins= library(palmerpenguins)
```

## Data Structures

### c()
The function c() combines or concatenates its arguments into a vector (a 1-dimensional data structure consisting of 1 or more elements).

* All of the elements must be of the same type
* I can't combine character and numeric types in the same call to c()

Here's two examples using numeric and character data types:

```{r}
## Create a vector of numbers:
num_vec  = c(1, 4, 8, 9, 13)

## Create a vector of characters:
char_vec = c("a", "fish", "data is cool")
```

You can show the contents of a vector by typing the name of the vector, or using the print() function.

```{r}
## Typing the name of the vector into the console prints the contents
num_vec
```

```{r}
## The print() function accomplishes the same task: 
print(char_vec)
```

### length()

The length() function allows you to see how many pieces are in a list or other object in your environment.


```{r}
length(char_vec)

length(num_vec)
```

### matrix()
The matrix() function allow you to create a matrix from a given set of values or a vector.

```{r}
matrix(num_vec)

matrix(1:8)
```

You can also specify the number of rows and columns in a matrix. 

```{r}
matrix(1:8, ncol=2, nrow=4)
```

R defaults to filling the matrix by columns, but you can change it to fill by row.

```{r}
matrix(1:8, ncol=2, nrow=4, byrow=TRUE)
```

### data.frame()
The data.frame() allows you to load data into R's environment as a collection of variables. 
A data frame can store many different types of data into 1 data frame, while a matrix is limited to only 1 type of data (usually numbers). 

```{r}
num_data= data.frame(num_vec)

char_data= data.frame(char_vec)
```
Now that the data is in a data frame you can view it as a collection, rather than a string of numbers.

```{r}
num_data
num_vec
```

### nrow()
The nrow() function produces the number of rows in the data frame (or matrix) you input.

```{r}
nrow(ginkgo)
```

### ncol()
Then ncol() function produes the number of columns in the data frame (or matrix) you input.

```{r}
ncol(ginkgo)
```

### dim()
The dim() function produces the dimensions of the data frame (or matrix) you input.

```{r}
dim(ginkgo)
```

## Subsetting

### $
Using the $ to subset allows you to pull out a specific named part of the data.

```{r}
width= ginkgo$max_width
```

### []
Using [] to subset allows you to pull out an object from the data.
To select the first row of the ginkgo data:

```{r}
ginkgo[1]
```

To select the element in row 2, column 3

```{r}
ginkgo[2,3]
```

To select the 3rd column of the ginkgo data

```{r}
ginkgo[,3]
```

### subset()
The subset() function allows you to pull out data based on a criteria that you set.
To retrieve all the data for Adelie penguins from the penguins dataset:
```{r}
dat_adelie=subset(penguins, species=="Adelie")
```

## Numerical Data Exploration

### summary()
The summary() function gives you an overview of what the data contains for all data types.
```{r}
summary(penguins)
```
From this summary you can see that there are 124 Gentoo penguins listed in the species column and 52 rows of data listed as Torgersen island.

### mean()
The mean() function calculates the mean of the data you specify. 

```{r}
mean(ginkgo$max_depth)
```
If you have a dataset that contains NAs, you must add in na.rm=TRUE
```{r}
mean(penguins$body_mass_g, na.rm=TRUE)
```

### sd()
The sd() function calculates the standard deviation of the data you specify.
```{r}
sd(ginkgo$max_depth)
```
If you have a dataset that contains NAs, you must add in na.rm=TRUE
```{r}
sd(penguins$body_mass_g, na.rm=TRUE)
```

## Graphical Data Exploration

### plot() 
The plot() function allows you to create a scatterplot of the data you specify.

```{r}
plot(max_width ~ max_depth, data=ginkgo)
```


You can assign many different arguments for the plot() function including:

* col=
Allows you to specify the color
* pch=
Allows you to change the shape of the plotting character
* cex=
Allows you to change the size of the plotting character
* main=
Allows you to label the plot (title)
* xlab=
Allows you to label the x-axis
* ylab=
Allows you to label the y-axis
* xlim=
Allows you to set the limits of the x-axis
* ylim=
Allows you to set the limits of the y-axis

```{r}
plot(max_width ~ max_depth, data=ginkgo, col="blue", pch=5,
     cex=2, main="Ginkgo Leaf Measurements",
     xlab= "Max Depth", ylab= "Max Width",
     xlim=c(0,160), ylim=c(0,140))
```

### hist()
This hist() function allows you to make a histogram from the data you specify.
```{r}
hist(penguins$flipper_length_mm, xlab="Flippger Length", main="Penguin Flipper Length mm")
```

* breaks=
Allows you to change the number of bars the histogram breaks the data into.

You can get a more detailed representation of the data with a greater number of breaks.

```{r}
hist(penguins$flipper_length_mm, breaks=30, xlab="Flippger Length", main="Penguin Flipper Length mm")
```


### boxplot()
The boxplot() function allows you to make a boxplot from the data you specify.

To create a simple boxplot of the Ginkgo petiole length:

```{r}
boxplot(ginkgo$petiole_length, xlab="petiole_length (mm)", main="Ginkgo Petiole Length")
```

To create a conditional boxplot showing the continuous variable of notch depth conditioned on the seeds_present column:

```{r}
boxplot(notch_depth ~ seeds_present, data=ginkgo, xlab="seeds_present", ylab= "notch_depth (mm)", main="Ginkgo Notch Depth Conditioned on Seed Presence")
```


### par(mfrow=)
The par() function combined with the mfrow= argument allows you to set the layout of the graphs you want to see at one time

Set mfrow=c(NUMBER OF ROWS, NUMBER OF COLUMNS) you want to have displayed

```{r}
par(mfrow= c(2,2))

hist(penguins$bill_length_mm, xlab="Bill Length (mm)", main="Penguin Bill Length",
     col="red")
     
hist(penguins$body_mass_g, breaks = 80, main = "Penguin Body Mass", 
     xlab = "Body Mass (g)", col="green")
     
hist(penguins$flipper_length_mm,
     ylim = c(0,50),
     main="Penguin Flipper Length",
     xlab="Flipper Length (mm)",
     col= "green")

hist(penguins$bill_depth_mm,
     main="Penguin Bill Depth",
     xlab="Bill Depth (mm)",
     col= "red")

dev.off()
```


## Distribution Functions

* dnorm(): the probability density, height of the curve

```{r}
dnorm(-1.96, mean=0, sd=1)
```

* pnorm()l: the cumulative probability density, area underneath the curve
```{r}
pnorm(-1.96, mean=0, sd=1)
```

* qnorm(): the quantile function

  You assign a probability and R gives you the boundary value

  If you want to find the 85th percentile of a normal distribution that has a mean of 70 and standard deviation of 3:

```{r}
qnorm(0.85,mean=70,sd=3)
```

* dbinom(): the probability mass function

  Allows you to ask: what is the probability that I observe a value of exactly x?

  Observing a count of exactly 3 successes, n=4, p=0.75
 
```{r}
 dbinom(3, 4, 0.75)
```

* pbinom(): the cumulative mass function

  Allows you to ask: what is the probability that I observe a value of x or less?

  observing a count of 3 or less, n=4, p=0.75 (3 or less includes 3)
  
```{r}
pbinom(3, 4, 0.75)
```

  observing more than 3, n=5, p=0.75 (more than 3 does not include 3)

```{r}
1- pbinom(3, 5, 0.75)
```

* qbinom(): the quantile function

  Allows you to ask: what is the median or 50th percentile?
  
  If you want to find the 10th quantile of a binomial distribution with 10 trials and probability of success on each trial = 0.4
  

```{r}
qbinom(.10, size=10, prob=.4)
```
